Diversity dynamics in Nymphalidae butterflies: Effect of phylogenetic uncertainty on diversification rate shift estimates

### 2014-11-11 MS_revised.tex
rm in .+ in

\documentclass[10pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{cite}
\usepackage{hyperref}
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

@@

In this study, the effect of phylogenetic uncertainty on the inferred
diversification splits by MEDUSA is amplified because some Nymphalidae
taxa appear to be strongly affected by long-branch attraction artifacts
\cite{pena2011}. Thus, the Bayesian runs are expected to recover alternative
topologies on the posterior distribution of trees, resulting in low
support and posterior probability values for the nodes. For example,
posterior probability values for clades in Satyrini are very low
\cite{wahlberg2009}. As a result, MEDUSA inferred a net diversification rate
increase for part of Satyrini in the maximum clade credibility tree, but
this was recovered in the MultiMEDUSA analysis in only in 13\% of the
random sample of trees.

@@

\section*{Conclusions}
We found that even though MEDUSA estimated several diversification
shifts in the maximum clade credibility tree of Nymphalidae, only a few
of these splits were found in more than 90\% of the trees from the
posterior distribution. In the literature, it is common practice that
conclusions are based on the splits estimated on the maximum clade
credibility tree. However, by using a MultiMEDUSA approach, we found
that for this Nymphalidae dataset some of these splits might be greatly
affected by phylogenetic uncertainty. Moreover, some of these splits can
be recovered either as increases or decreases in net diversification
rate depending on the tree from the posterior distribution that was used
for analysis. This means that contradictory conclusions would be made if
only the maximum clade credibility tree was used for analysis. We
recommend that all datasets should be analyzed using both approaches,
MEDUSA and MultiMEDUSA, in order to test whether the results are robust
when phylogenetic uncertainty is taken into account.
Moreover, we also found that the results from BAMM, without taking into account
phylogenetic uncertainty, correspond better to those of MultiMEDUSA than MEDUSA.

>>>>

\documentclass[10pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{cite}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{lineno}
\linenumbers
\DisableLigatures[f]{encoding = *, family = * }

@@

In this study, the effect of phylogenetic uncertainty on the inferred
diversification splits by MEDUSA is amplified because some Nymphalidae
taxa appear to be strongly affected by long-branch attraction artifacts
\cite{pena2011}. Thus, the Bayesian runs are expected to recover alternative
topologies on the posterior distribution of trees, resulting in low
support and posterior probability values for the nodes. For example,
posterior probability values for clades in Satyrini are very low
\cite{wahlberg2009}. As a result, MEDUSA inferred a net diversification rate
increase for part of Satyrini in the maximum clade credibility tree, but
this was recovered in the MultiMEDUSA analysis in only 13\% of the
random sample of trees.

@@

\section*{Conclusions}
We found that even though MEDUSA estimated several diversification
shifts in the maximum clade credibility tree of Nymphalidae, only a few
of these splits were found in more than 90\% of the trees from the
posterior distribution. In the literature, it is common practice that
conclusions are based on the splits estimated on the maximum clade
credibility tree. However, by using a MultiMEDUSA approach, we found
that for this Nymphalidae dataset some of these splits might be greatly
affected by phylogenetic uncertainty. Moreover, some of these splits can
be recovered either as increases or decreases in net diversification
rate depending on the tree from the posterior distribution that was used
for analysis. This means that contradictory conclusions would be made if
only the maximum clade credibility tree was used for analysis. We
recommend that all datasets should be analyzed using both approaches,
MEDUSA and MultiMEDUSA, in order to test whether the results are robust
when phylogenetic uncertainty is taken into account.
The results from BAMM recovered only two significant shifts in our MCC tree
as this method is more conservative than the MEDUSA approach. One of these 
shifts is the Satyrini clade, which was recovered by MEDUSA in a few trees 
from the posterior distribution. Thus, it might be important to
take into account phylogenetic uncertainty in BAMM as well.
